---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# speedyseq

The goal of `speedyseq` is to accelerate common operations that are currently
very slow in `phyloseq`: the function `psmelt` (and the plotting functions that
use it) and the taxonomic aggregation functions `tax_glom` and `tip_glom`.

The current version of `speedyseq` reimplements `psmelt()` to be much faster
than `phyloseq`'s current version, and includes copies of `plot_bar()`,
`plot_heatmap()`, and `plot_tree()` so that when called from `speedyseq` the
faster `psmelt()` will be used. It also implements a faster version of
`tax_glom()`. 

My aim is for these functions to create output that is identical
to what would be obtained from `phyloseq`; however, I have not tested this
extensively, and offer no gaurantee that these versions won't fail in obvious
or subtle ways. If you have any problems or find any discrepancies from
`phyloseq`'s behavior, please post an issue.

## Installation

Install with `devtools`
```r
# install.packages("devtools")
devtools::install_github("mikemc/speedyseq")
```

## Usage

Method 1: Call `speedyseq` functions explicitly when you want to use
`speedyseq`'s version instead of `phyloseq`:
```{r}
library(phyloseq)
data(GlobalPatterns)
system.time(
    # Calls phyloseq's psmelt
    df1 <- psmelt(GlobalPatterns) # slow
)
system.time(
    df2 <- speedyseq::psmelt(GlobalPatterns) # fast
)
```

Method 2: Load `speedyseq` after `phyloseq`, which will cause calls to the
overlapping function names to go to `speedyseq` by default:
```{r}
library(phyloseq)
library(speedyseq)
data(GlobalPatterns)
system.time(
    ps1 <- phyloseq::tax_glom(GlobalPatterns, "Genus") # slow
)
system.time(
    # Calls speedyseq's tax_glom
    ps2 <- tax_glom(GlobalPatterns, "Genus") # fast
)
all.equal(taxa_names(ps1), taxa_names(ps2))
all.equal(otu_table(ps1), otu_table(ps2))
all.equal(tax_table(ps1), tax_table(ps2))
all.equal(phy_tree(ps1), phy_tree(ps2))
```
